Topological Equivariant Artist Model for Visualization Library Architecture

### 2024-12-27 paper/main.tex
a presheaf is a presheaf

\begin{definition}\label{def:atct:sheaf}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
\begin{itemize}
  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values $\dsection^{a}\vert_{\openset_i} =  \dsection^{b}\vert_{\openset_i}$ over the open cover $\bigcup_{i\in I} \openset_{i} \subset \openset$ (indexed by $I$).
  \item \textit{gluing} the union of sections defined over subspaces $\dsection^{i} \in \Gamma(\openset_i, \dtotal|_{openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}\vert_{\openset_i\cap\openset_j} =  \dsection^{j}\vert_{\openset_i\cap\openset_j}$
  \end{itemize}
\end{definition}

@@

\subsection{Homeomorphism}
As mentioned in \autoref{sec:related-work:continuity}, preserving the topology of a visualization means that each discrete piece of differentiable visual information corresponds to a distance element of the dataset\cite{ziemkiewiczEmbeddingInformationVisualization2009} in a way where the organization of elements is preserved. A generalization of this condition is the idea that the graphic space can be collapsed into the data indexing space, which means that the data base space is a deformation retraction of the graphic base space\cite{hatcherAlgebraicTopology2002}. By defining the indexing look up function \vindex, introduced in \autoref{sec:atct:sheaves}, to be
\begin{equation}
  \vindex: \dbase \times I \rightarrow \dbase s.t \vindex(\dbasepoint) = \dbasepoint \forall \dbasepoint \in \gbase
\end{equation}

@@

Because the artist is a functor of sheaves, the artist preserves the inclusion and restriction conditions of a presheaf \autoref{def:atct:presheaf} and the locality and gluing conditions of a sheaf \autoref{def:atct:sheaf} when mapping between subspaces $\vindex(\opensetg) \rightarrow \openset$. This bookkeeping is necessary for any visualization technique that selectively acts on different pieces of a data set; for example streaming visualizations \cite{krstajicVisualizationStreamingData2013} and panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}

\subsection{Equivariance}
\label{sec:artist:equiv}
As introduced in \autoref{sec:related-work:equivariance}, data and the corresponding visual encoding are expected to have compatible structure. This structure can be formally expressed as actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$. We generalize from binary operations to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator. We describe the changes on the graphic side as changes in measurements \measure\, which are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic. The visual variables \cite{bertinIIPropertiesGraphic2011} are a subset of measurable components. For example, a measurement of a scatter marker could be its color (e.g. red) or its x position (e.g. 5).

>>>>

\begin{definition}\label{def:atct:sheaf}\cite{bakerEuclideanSpaceMathsSheaf,spanier1989algebraic} A \textbf{sheaf} is a presheaf that satisfies the following two axioms
\begin{itemize}
  \item \textit{locality} two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values $\dsection^{a}\vert_{\openset_i} =  \dsection^{b}\vert_{\openset_i}$ over the open cover $\bigcup_{i\in I} \openset_{i} \subset \openset$ (indexed by $I$).
  \item \textit{gluing} the union of sections defined over subspaces $\dsection^{i} \in \Gamma(\openset_i, \dtotal|_{\openset_i})$ is equivalent to a section defined over the whole space $\dsection\vert_{\openset_i} = \dsection^{i}$ for all $i\in I$ if all pairs of sections agree on overlaps $\dsection^{i}\vert_{\openset_i\cap\openset_j} =  \dsection^{j}\vert_{\openset_i\cap\openset_j}$
  \end{itemize}
\end{definition}

@@

\subsection{Homeomorphism}
As mentioned in \autoref{sec:related-work:continuity}, preserving the topology of a visualization means that each discrete piece of differentiable visual information corresponds to a dist element of the dataset\cite{ziemkiewiczEmbeddingInformationVisualization2009} in a way where the organization of elements is preserved. A generalization of this condition is the idea that the graphic space can be collapsed into the data indexing space, which means that the data base space is a deformation retraction of the graphic base space\cite{hatcherAlgebraicTopology2002}. By defining the indexing look up function \vindex, introduced in \autoref{sec:atct:sheaves}, to be
\begin{equation}
  \vindex: \dbase \times I \rightarrow \dbase\;s.t\;\vindex(\dbasepoint) = \dbasepoint\;\forall \dbasepoint \in \gbase
\end{equation}

@@

When the data and graphic containers are implemented as sheaves, the artist preserves the sheaf constraints because it is using the sheaf constraint satisfying map $\vindex$ to maintain the correspondence between data and graphing indexing spaces. Given a subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$, all the data in the larger index is included in the smaller set $\Gamma(\dbase_1, \dtotal) \xrightarrow{i^*} \Gamma(\dbase_2, \dtotal)$ such that visualization specs of the smaller set must include specs for the larger set of data values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$. While the curves in \autoref{fig:atct:presheaf} are meant as stand-ins for the data, they also serve to illustrate that the relationship is identical in the graphic space such that there may be more graphics available for data from a smaller indexing space because there may be more values on the smaller indexing space. The artist also preserves the \textit{locality} axiom in that identical datasets $\dsection^{a} = \dsection^{b}$ generate identical specifications $\vartist(\dsection^a) = \vartist(\dsection^b)$. The artist also respects \textit{gluing} in that graphics are composable, as described in \autoref{sec:artist:operators}, such that a set of artists that generate graphics for parts of a dataset are equivalent to an artist that generates a graphic for the whole dataset. Managing subsets of data and on demand data in a structure preserving way is necessary to support visualizing streaming data \cite{krstajicVisualizationStreamingData2013} and techniques that rely on shifting or resampling the indexing space, such as  panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}.

\note{stop here}
\subsection{Equivariance}
\label{sec:artist:equiv}
As introduced in \autoref{sec:related-work:equivariance}, data and the corresponding visual encoding are expected to have compatible structure. This structure can be formally expressed as actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$. We generalize from binary operations to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator. We describe the changes on the graphic side as changes in measurements \measure\, which are scaler or vector components of the rendered graphic that can be quantified, such as the color, position, shape, texture, or rotation angle of the graphic. The visual variables \cite{bertinIIPropertiesGraphic2011} are a subset of measurable components. For example, a measurement of a scatter marker could be its color (e.g. red) or its x position (e.g. 5).
